---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
# Create the deaths dataset in long format
deaths <- av %>%
  select(URL, Name.Alias, Death1, Death2, Death3, Death4, Death5) %>%
  pivot_longer(
    cols = starts_with("Death"),
    names_to = "Time",
    values_to = "Death"
  ) %>%
  mutate(
    Time = readr::parse_number(Time)
  )

# Display the first few rows of the deaths dataset
head(deaths)
```

## Restructuring Returns Data to Long Format

Similarly, let's reshape the Return columns into a long format dataset:

```{r}
# Create the returns dataset in long format
returns <- av %>%
  select(URL, Name.Alias, Return1, Return2, Return3, Return4, Return5) %>%
  pivot_longer(
    cols = starts_with("Return"),
    names_to = "Time",
    values_to = "Return"
  ) %>%
  mutate(
    Time = readr::parse_number(Time)
  )

# Display the first few rows of the returns dataset
head(returns)
```

## Calculating Average Number of Deaths per Avenger

Now, let's calculate the average number of deaths an Avenger suffers, correctly accounting for those with zero deaths:

```{r}
# Count deaths for each character
death_counts <- deaths %>%
  filter(Death == "YES") %>%
  group_by(URL, Name.Alias) %>%
  summarise(death_count = n(), .groups = "drop")

# Get the full list of all Avengers
all_avengers <- av %>%
  select(URL, Name.Alias) %>%
  distinct()

# Join the death counts with all avengers to include those with zero deaths
complete_death_counts <- all_avengers %>%
  left_join(death_counts, by = c("URL", "Name.Alias")) %>%
  mutate(death_count = replace_na(death_count, 0))

# Calculate the average number of deaths per Avenger
avg_deaths <- mean(complete_death_counts$death_count)

# Print the result
cat("The average number of deaths per Avenger is:", round(avg_deaths, 2))
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.


### Harsh's Work:

### FiveThirtyEight Statement

"Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team. Let’s put it this way: If you fall from four or five stories up, there’s a 50 percent chance you die. Getting a membership card in the Avengers is roughly like jumping off a four-story building"

### Include the code

```{r}
# Count the number of Avengers with each number of deaths
death_distribution <- complete_death_counts %>%
  count(death_count) %>%
  mutate(percentage = n / sum(n) * 100)

# Display the distribution
death_distribution
```
### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

60% of the Avengers haven't died once, directly oposing what they have said that 69% have died at least once.

### Aiden's Work:

### FiveThirtyEight Statement

"I counted 89 total deaths — some unlucky Avengers are basically Meat Loaf with an E-ZPass — and on 57 occasions the individual made a comeback. Maybe they didn’t actually die and were secretly in the Microverse, or they stayed on Franklin Richards‘s or the Scarlet Witch‘s good side in life, or they were dragged back into Avenging by the Chaos King or the Grim Reaper, or perhaps a colleague made a deal with time travelers. Who knows!"

"But you can only tempt death so many times. There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.8"

### Include the code

```{r}
#Create a list of all the characters with at least one death
first_three_death_list <- av %>%
  select(URL, Name.Alias, Death1, Return1, Death2, Return2, Death3, Return3) %>%
  filter(Death1 == "YES")

#Filter out all the people that don't return after their first death
no_return_first_death <- first_three_death_list %>%
  filter(Return1 == "NO") %>%
  nrow()

#Calculate how many of the characters return after their first death
return_first_death_percent <- (1 - no_return_first_death / nrow(first_three_death_list)) * 100

#Characters who died twice
has_second_death <- first_three_death_list %>%
  filter(Death2 == "YES")

#Characters who died 3 times
has_third_death <- first_three_death_list %>%
  filter(Death3 == "YES")

#Number of characters who returned from their second death
second_return <- has_second_death %>%
  filter(Return2 == "YES") %>%
  nrow()

#Number of characters who returned from their third death
third_return <- has_third_death %>%
  filter(Return3 == "YES") %>%
  nrow()

#Number of characters who returned from a 2nd and/or 3rd death divided by the number of characters who had a 2nd or 3rd death
return_later_percent <- (second_return + third_return) / ((nrow(has_second_death) + nrow(has_third_death)))
```
### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

The first statement that there is a 2 in 3 chance that characters return after their 3rd death checked out, my data also got 66.66% of characters returning. The second statement that there is only a 50% chance of recovery from a 2nd or 3rd death was also accurate in my findings, with exactly 50% of 2nd or 3rd deaths being returned from. Overall in this section they were accurate based on my findings. 
