---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
# Create the deaths dataset in long format
deaths <- av %>%
  select(URL, Name.Alias, Death1, Death2, Death3, Death4, Death5) %>%
  pivot_longer(
    cols = starts_with("Death"),
    names_to = "Time",
    values_to = "Death"
  ) %>%
  mutate(
    Time = readr::parse_number(Time)
  )

# Display the first few rows of the deaths dataset
head(deaths)
```

## Restructuring Returns Data to Long Format

Similarly, let's reshape the Return columns into a long format dataset:

```{r}
# Create the returns dataset in long format
returns <- av %>%
  select(URL, Name.Alias, Return1, Return2, Return3, Return4, Return5) %>%
  pivot_longer(
    cols = starts_with("Return"),
    names_to = "Time",
    values_to = "Return"
  ) %>%
  mutate(
    Time = readr::parse_number(Time)
  )

# Display the first few rows of the returns dataset
head(returns)
```

## Calculating Average Number of Deaths per Avenger

Now, let's calculate the average number of deaths an Avenger suffers, correctly accounting for those with zero deaths:

```{r}
# Count deaths for each character
death_counts <- deaths %>%
  filter(Death == "YES") %>%
  group_by(URL, Name.Alias) %>%
  summarise(death_count = n(), .groups = "drop")

# Get the full list of all Avengers
all_avengers <- av %>%
  select(URL, Name.Alias) %>%
  distinct()

# Join the death counts with all avengers to include those with zero deaths
complete_death_counts <- all_avengers %>%
  left_join(death_counts, by = c("URL", "Name.Alias")) %>%
  mutate(death_count = replace_na(death_count, 0))

# Calculate the average number of deaths per Avenger
avg_deaths <- mean(complete_death_counts$death_count)

# Print the result
cat("The average number of deaths per Avenger is:", round(avg_deaths, 2))
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.


### Harsh's Work:

### FiveThirtyEight Statement

"Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team. Let’s put it this way: If you fall from four or five stories up, there’s a 50 percent chance you die. Getting a membership card in the Avengers is roughly like jumping off a four-story building"

### Include the code

```{r}
# Count the number of Avengers with each number of deaths
death_distribution <- complete_death_counts %>%
  count(death_count) %>%
  mutate(percentage = n / sum(n) * 100)

# Display the distribution
death_distribution
```
### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

60% of the Avengers haven't died once, directly oposing what they have said that 69% have died at least once.

Upload your changes to the repository. Discuss and refine answers as a team.
